Ornaments as indicators of social changes resulting from indirect effects of colonialism in northeastern Taiwan

### 2020-07-13 analysis/paper/paper.Rmd
changed title

---
title: "Ornaments as indicators of social changes in northeastern Taiwan before and after the European colonial period"
author:
  - Li-Ying Wang
  - Ben Marwick
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      #reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/antiquity.csl" # Insert path for the bib-style
abstract: |
  Long-lasting indirect impacts on Indigenous peoples in the periphery of colonial control are poorly understood, especially in East Asia. Trade ornaments from Kiwulan (1350-1850 AD) in northeastern Taiwan show the indirect impacts of European colonial activities on local societies. The diversity of ornaments was greater during the period of European presence compared to previous periods, and their spatial distribution was more clustered. This hints at increasing social inequality resulting from colonial influence. Ornaments give insights into the increasing social inequality stimulated by the European colonial presence, and show the agency of Indigenous people to incorporate ornaments into their social system.
keywords: |
  Historical archaeology; European colonial period; 17th century; Trade ornaments; Exchange networks; spatial pattern; East Asia
highlights: |
  1. Evidence of the quantity, variety, and spatial distribution of trade ornaments in both local cultural and colonial context is provided.
  2. Results of hypothesis testing point to greater amount and diversity of ornament types with non-random, and most highly concentrated spatial clusters during the European presence.
  3. Trade ornament as a prestige good may result in an increase in social inequality due to competition of wealth accumulation between individuals.
  4. Trade ornament can be a good proxy to detect indirect colonial influence on local Indigenous populations and inform on social changes relating to trade with colonial groups.
---

@@

Studying change over time at a site like Kiwulan has some distinct challenges. First, the archaeological deposits are relatively thin. This makes it difficult for excavators to clearly identify distinct chronological phase boundaries. In thin deposits, site formation processes can quickly blur the boundaries or completely mix deposits from different time periods. Second, the time periods we aim to compare are relatively short, just a few human generations. This limits the usefulness of common chronometric methods such as radiocarbon dating, where the age errors are relatively large compared to the periods we are studying. Most of the previously obtained radiocarbon ages from Kiwulan do not span the time period we are investigating here (Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data)). Third, pericolonial contexts such as Kiwulan can involve non-Western notions and experiences of time. For example, a Western concept of time seeks clear temporal divides, and may create material signature of this, but Indigenous experiences of time may prioritize continuity and continuums of gradual change, and their material culture traces may be reflect that intentional experience of time [@Trabert2018; @Scheiber2012]. Despite the relatively short period of the European colonization in Taiwan (38-years-long Dutch occupation and 16-years-long Spanish occupation), the indirect impacts of the Europeans through trade networks could go beyond these temporal boundaries and last longer. This is because the use of foreign goods in many Indigenous societies often goes through processes of negotiations resistance, and transformation. A static time frame may not capture the processes well [@Scheiber2012]. Fourth, the circulation of foreign goods that we use as chronological indicators might start earlier than the historically documented European presence in Taiwan. Similarly, the use of the foreign goods in Indigenous societies could continue after the end of the European colonization [@Mitchell2010]. This limits the usefulness of specific artifact types as chronological markers, and implies they have an error range similar to radiocarbon ages. Unfortunately, we do not have a way to accurately estimate these errors at Kiwulan. Ideally, we would divide the deposits at Kiwulan into many time slices to investigate the possibility of a gradual versus sudden change, and assess errors in the accuracy of artefacts as temporal markers. However, this would result in many analytical units with no finds, so we have grouped excavation units together to create minimum sample sizes suitable for addressing our research questions. The details about the distribution of the temporal indicators, accompanied with stratigraphy data, radiocarbon dates, and archaeological contexts for each sampled excavation unit are provided in Supplementary Online Materials.

>>>>

---
title: "Ornaments as indicators of social changes resulting from indirect effects of colonialism in northeastern Taiwan"
author:
  - Li-Ying Wang
  - Ben Marwick
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      #reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/antiquity.csl" # Insert path for the bib-style
abstract: |
  Long-lasting indirect impacts on Indigenous peoples in the periphery of colonial control are poorly understood, especially in East Asia. Trade ornaments from Kiwulan (1350-1850 AD) in northeastern Taiwan show the indirect impacts of European colonial activities on local societies. The diversity of ornaments was greater during the period of European presence compared to previous periods, and their spatial distribution was more clustered. This hints at increasing social inequality resulting from colonial influence. Ornaments give insights into the increasing social inequality stimulated by the European colonial presence, and show the agency of Indigenous people to incorporate ornaments into their social system.
keywords: |
  Historical archaeology; European colonial period; 17th century; Trade ornaments; Exchange networks; spatial pattern; East Asia
highlights: |
  1. Evidence of the quantity, variety, and spatial distribution of trade ornaments in both local cultural and colonial context is provided.
  2. Results of hypothesis testing point to greater amount and diversity of ornament types with non-random, and most highly concentrated spatial clusters during the European presence.
  3. Trade ornament as a prestige good may result in an increase in social inequality due to competition of wealth accumulation between individuals.
  4. Trade ornament can be a good proxy to detect indirect colonial influence on local Indigenous populations and inform on social changes relating to trade with colonial groups.
---

@@

Studying change over time at a site like Kiwulan has some distinct challenges. First, the archaeological deposits are relatively thin. This makes it difficult for excavators to clearly identify distinct chronological phase boundaries. In thin deposits, site formation processes can quickly blur the boundaries or completely mix deposits from different time periods. Second, the time periods we aim to compare are relatively short, just a few human generations. This limits the usefulness of common chronometric methods such as radiocarbon dating, where the age errors are relatively large compared to the periods we are studying. Most of the previously obtained radiocarbon ages from Kiwulan do not span the time period we are investigating here (Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data)). Third, pericolonial contexts such as Kiwulan can involve non-Western notions and experiences of time. For example, a Western concept of time seeks clear temporal divides, and may create material signature of this, but Indigenous experiences of time may prioritize continuity and continuums of gradual change, and their material culture traces may reflect that intentional experience of time [@Trabert2018; @Scheiber2012]. Despite the relatively short period of the European colonization in Taiwan (38-years-long Dutch occupation and 16-years-long Spanish occupation), the indirect impacts of the Europeans through trade networks could go beyond these temporal boundaries and last longer. This is because the use of foreign goods in many Indigenous societies often goes through processes of negotiations, resistance, and transformation. A static time frame may not capture the processes well [@Scheiber2012]. Fourth, the circulation of foreign goods that we use as chronological indicators might start earlier than the historically documented European presence in Taiwan. Similarly, the use of the foreign goods in Indigenous societies could continue after the end of the European colonization [@Mitchell2010]. This limits the usefulness of specific artifact types as chronological markers, and implies they have an error range similar to radiocarbon ages. Unfortunately, we do not have a way to accurately estimate these errors at Kiwulan. Ideally, we would divide the deposits at Kiwulan into many time slices to investigate the possibility of a gradual versus sudden change, and assess errors in the accuracy of artefacts as temporal markers. However, this would result in many analytical units with no finds, so we have grouped excavation units together to create minimum sample sizes suitable for addressing our research questions. The details about the distribution of the temporal indicators, accompanied with stratigraphy data, radiocarbon dates, and archaeological contexts for each sampled excavation unit are provided in Supplementary Online Materials.
